---
title: "Group 26 CA pollution burden"
author: "Taylor Gries and Cristina Robinson"
date: "11/12/2021"
output: html_document
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) # Attach packages
library(tidyverse)
library(janitor)
library(here)
library(hexbin) # Make sure to install "hexbin" in console
```

```{r}
ces <- read_csv(here("data", "ces3demographics.csv")) %>%
  clean_names()
```

```{r}
ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) + 
geom_point() + # create basic scatterplot graph 
geom_hex() + # create 2D hexbin graph 
  labs(x = "% White", # add in axis labels
       y = "CalEnviroScreen Score") +
  scale_fill_gradient(low = "cyan", high = "darkcyan") + # customize color gradient
  theme_minimal() # update graph to theme minimal 

```

**Figure 1**: The CalEnviroScreen pollution burden score for % of white populations based on the California census.

It is observed in Figure 1 that higher percent of white populations (and with higher counts) have a lower CalEnviroScreen Score, which is correlated to a lower cumulative pollution burden and lower vulnerable population characteristics. The CalEnviroScreen 3.0 indicators include pollution burdens such as exposures and environmental effects, as well as population characteristics such as sensitive populations and socioeconomic factors, resulting in the final CalEnviroScreen 3.0 Score.

**Data Citation**: California Office of Environmental Health Hazard Assessment. Wieland, W. 2018. "CalEnviroScreen 3.0: A Tool For Advancing Environmental Justice", California Environmental Protection Agency. https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30